---
title: "ass2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


read data

```{r}
conn<-read.csv("D:\\SSS\\MSC\\dst\\data\\ass2\\MAC.csv",header = T)
conn
```


confirm is there any missingness in each feature and find the proportions of missingness for each protocol

```{r}
# there is missingness in duration
mtab1=data.frame(
    missingduration=is.na(conn[,"duration"]),
    proto=conn[,"proto"])
mtab1=table(mtab1)
(apply(mtab1,2,function(x)x/sum(x)))
```


```{r}
# there is missingness in orig_bytes
mtab2=data.frame(
    missingorigbytes=is.na(conn[,"orig_bytes"]),
    proto=conn[,"proto"])
mtab2=table(mtab2)
(apply(mtab2,2,function(x)x/sum(x)))
```


```{r}
# there is missingness in resp_bytes
mtab3=data.frame(
    missingrespbytes=is.na(conn[,"resp_bytes"]),
    proto=conn[,"proto"])
mtab3=table(mtab3)
(apply(mtab3,2,function(x)x/sum(x)))
```


```{r}
# the whole local_orig is NA
mtab4=data.frame(
    missinglocalorig=is.na(conn[,"local_orig"]),
    proto=conn[,"proto"])
mtab4=table(mtab4)
mtab4
```

```{r}
# no missing values in missed_bytes
mtab5=data.frame(
    missingmissedbytes=is.na(conn[,"missed_bytes"]),
    proto=conn[,"proto"])
mtab5=table(mtab5)
mtab5
```

```{r}
# no missing values in history
mtab6=data.frame(
    missinghistory=is.na(conn[,"history"]),
    proto=conn[,"proto"])
mtab6=table(mtab6)
mtab6
```

```{r}
# then table the missed_bytes, find that there is only one value--0
mb=table(conn$missed_bytes)
mb
#hi=table(conn$history)
#hi
```


```{r}
#conn_state=table(conn$conn_state)
#conn_state
```


```{r}
# no missing values in orig_pkts
mtab7=data.frame(
    missingorigpkts=is.na(conn[,"orig_pkts"]),
    proto=conn[,"proto"])
mtab7=table(mtab7)
mtab7
```


```{r}
# no missing values in service
mtab8=data.frame(
    missingservice=is.na(conn[,"service"]),  
    proto=conn[,"proto"])
mtab8=table(mtab8)
mtab8
```

planned to remove the outliers, but once DBSCAN is used, there is no need to do it i guess?

```{r}
#thist=hist(conn[,"duration"],breaks=101,plot=FALSE)
#plot(thist$mids,thist$density,log="y",type="b",
     #xlab="duration",ylab="histogram density")
#threshold=1200
#abline(v=threshold,col=2) 
```

```{r}
#obhist=hist(conn[,"orig_bytes"],breaks=101,plot=FALSE)
#plot(obhist$mids,obhist$density,log="y",type="b",
     #xlab="orig_bytes",ylab="histogram density")
##threshold=1100
##abline(v=threshold,col=2) 
```


```{r}
#rbhist=hist(conn[,"resp_bytes"],breaks=101,plot=FALSE)
#plot(rbhist$mids,rbhist$density,log="y",type="b",
     #xlab="orig_bytes",ylab="histogram density")
#threshold=1.5e+06
#abline(v=threshold,col=2) 
```


```{r}
#cshist=hist(conn[,"orig_ip_bytes"],breaks=101,plot=FALSE)
#plot(cshist$mids,cshist$density,log="y",type="b",
     #xlab="orig_ip_bytes",ylab="histogram density")
#threshold=1e+06
#abline(v=threshold,col=2) 
```


then drop some columns which are useless and may negatively affect the genelization ability.

```{r}
conn<-conn[,-(1:3)]
conn
```

```{r}
conn<-conn[,-(1:6)]
conn
```
```{r}
conn<-conn[,-(4:7)]
conn
```

then work with svds

```{r}
conncols=c("orig_ip_bytes","resp_ip_bytes","orig_pkts","resp_pkts")

conn_all_scaled <- apply(conn[,conncols], 2, scale)
indices=!is.na(conn[,"duration"]) #the rows of NA in other two columns is the same
completedata=conn[indices,]
conn_all.svd=svd(conn_all_scaled[indices,])
```


```{r}
#conn_all.svd
```

plot the data, then we can find that we need first 3 PCs, thus nPCs=3
```{r}
plot(conn_all.svd$d,xlab="Eigenvalue index",ylab="Eigenvalue",log="y")
```

parameter tuning for DBSCAN

find suitable eps parameter using a k-NN plot for k = dim + 1 to Look for the knee

```{r}
library("dbscan")
test=kNNdist(conn_all.svd$u[,1:3], k = 5,all=TRUE)
testmin=apply(test,1,min)
```

test what sort of thresholds for the parameter "eps" might be useful

```{r}
plot(sort(testmin[testmin>1e-8]),log="y")
threshholds= c(0.01,0.001,0.0001,0.00001,0.000001)
abline(h=c(0.01,0.001,0.0001,0.00001,0.000001))
abline(h=0.0001,col="red")  ##choose this
```

do the clustering with different minPts


```{r}
dbscanres5=dbscan(conn_all.svd$u[,1:3],0.0001,5)
dbscanres5
```

```{r}
dbscanres10=dbscan(conn_all.svd$u[,1:3],0.0001,10)
dbscanres10
```

```{r}
dbscanres20=dbscan(conn_all.svd$u[,1:3],0.0001,20)
dbscanres20
```

```{r}
dbscanres30=dbscan(conn_all.svd$u[,1:3],0.0001,30)
dbscanres30
```

```{r}
dbscanres50=dbscan(conn_all.svd$u[,1:3],0.0001,50)
dbscanres50
```

```{r}
dbscanres70=dbscan(conn_all.svd$u[,1:3],0.0001,70)
dbscanres70
```

```{r}
dbscanres100=dbscan(conn_all.svd$u[,1:3],0.0001,100)
dbscanres100
```

```{r}
dbscanres150=dbscan(conn_all.svd$u[,1:3],0.0001,150)
dbscanres150
```

```{r}
dbscanres200=dbscan(conn_all.svd$u[,1:3],0.0001,200)
dbscanres200
```

```{r}
dbscanres300=dbscan(conn_all.svd$u[,1:3],0.0001,300)
dbscanres300
```

```{r}
dbscanres400=dbscan(conn_all.svd$u[,1:3],0.0001,400)
dbscanres400
```

plot the results of clusters

```{r}
pairs(conn_all.svd$u[,1:3], col = dbscanres5$cluster + 1L)
```

```{r}
pairs(conn_all.svd$u[,1:3], col = dbscanres10$cluster + 1L)
```

```{r}
pairs(conn_all.svd$u[,1:3], col = dbscanres20$cluster + 1L)
```

```{r}
pairs(conn_all.svd$u[,1:3], col = dbscanres30$cluster + 1L)
```

```{r}
pairs(conn_all.svd$u[,1:3], col = dbscanres50$cluster + 1L)
```

```{r}
pairs(conn_all.svd$u[,1:3], col = dbscanres70$cluster + 1L)
```

```{r}
pairs(conn_all.svd$u[,1:3], col = dbscanres100$cluster + 1L)
```

```{r}
pairs(conn_all.svd$u[,1:3], col = dbscanres150$cluster + 1L)
```

```{r}
pairs(conn_all.svd$u[,1:3], col = dbscanres200$cluster + 1L)
```

```{r}
pairs(conn_all.svd$u[,1:3], col = dbscanres300$cluster + 1L)
```

```{r}
pairs(conn_all.svd$u[,1:3], col = dbscanres400$cluster + 1L)
```

then choose the minPts=20 to get 13 clusters which are close to Alex's results.

plot the results in which minPts=20

```{r}
png(file = "conndbscan20.png")
op <- par(mfrow=c(3,1))
    plot(conn_all.svd$u[,1],
         conn_all.svd$u[,2],xlab="",
         ylab="",main=paste("K=41 + noise"),
         col=c("#66666666",rainbow(41))[dbscanres20$cluster+1],pch=19,cex=5)
    plot(conn_all.svd$u[,1],
         conn_all.svd$u[,3],xlab="",
         ylab="",main=paste("K=41 + noise"),
         col=c("#66666666",rainbow(41))[dbscanres20$cluster+1],pch=19,cex=5)
    plot(conn_all.svd$u[,2],
         conn_all.svd$u[,3],xlab="",
         ylab="",main=paste("K=41 + noise"),
         col=c("#66666666",rainbow(41))[dbscanres20$cluster+1],pch=19,cex=5)
par(op)
dev.off()
```



```{r}
for (i in 0:13) {
  dfi<-conn[dbscanres20$cluster==i,]
}
dfi
```


```{r}
for (i in 0:13) {
  dfi[is.na(dfi[,"duration"]),"duration"]=mean(!is.na(dfi[,"duration"]))
 dfi[is.na(dfi[,"orig_bytes"]),"orig_bytes"]=mean(!is.na(dfi[,"orig_bytes"]))
  dfi[is.na(dfi[,"resp_bytes"]),"resp_bytes"]=mean(!is.na(dfi[,"resp_bytes"]))
}
dfi

```



```{r}
#install.packages("Hmisc")
#library(Hmisc)
```













